#Activity Monitoring Dataset
###Data consists of number of steps recorded at 5 min interval for a 2 month period


```{r}
setwd("C:\\Users\\Hasan\\Statistics\\R\\rep data activity")
data=read.csv("activity.csv")
data$date=as.Date(data$date)
head(data)
```


**Calculting the total number of steps taken each day**
```{r message=FALSE ,warning=FALSE}
library(dplyr)
sum.day=data %>%
  group_by(date) %>%
                summarise(sum=sum(steps,na.rm=TRUE))
head(sum.day)
```


**Histogram of the total number of steps taken each day**
```{r}
breaks=with(sum.day,seq(min(sum),max(sum),length=10))
with(sum.day,hist(sum,breaks=breaks,xaxt="n",ylim=c(0,20),xlab="sum of steps",
     main="Histogram of the total number of steps taken each day",col=rgb(0,0,0,0.35)))
axis(1,at=breaks,labels=ceiling(breaks),las=2,cex.axis=0.75)
box()
```

```{r}
mean.day=mean(sum.day$sum)
print(mean.day)
median.day=median(sum.day$sum)
print(median.day)
```

**Calculting the average number of steps taken in each interval in 2 months**
```{r}
sum.int=data %>%
  group_by(interval) %>%
  summarise(mean=mean(steps,na.rm=TRUE))
head(sum.int)
```
**Time series plot of the average number of steps taken in each interval**
```{r}
with(sum.int,plot(interval,mean,type="l",
                  xlab="5 minute intervals",ylab="mean steps per interval",
                  main="time series plot of the number of steps taken in each interval"))
```

**The 5 min interval having the maximum number of steps on an average**
```{r}
sum.int[which.max(sum.int$mean),"interval"]
```

**Imputation**
The impuation of missing values is carried out by taking a random sample from the available data and putting them in place of the missing values
This incorporates a randomness in the data and provies greated precision
```{r}
num.na=sum(is.na(data$steps))
pos.na=which(is.na(data$steps))
impute=sample(data$steps[-pos.na],num.na)
data.imp=data
data.imp[pos.na,"steps"]=impute
head(data.imp)
```

**Calculating the total number of steps taken each day after imputation**
```{r}
sum.day.imp=data.imp %>%
  group_by(date) %>%
  summarise(sum=sum(steps,na.rm=TRUE))
head(sum.day.imp)
```
**Histogram of total number of steps taken each day before and after imputation**
```{r}
par(mfrow=c(1,2),oma=c(0,0,2,0))
breaks=with(sum.day,seq(min(sum),max(sum),length=10))

with(sum.day,hist(sum,breaks=breaks,freq=TRUE,xaxt="n",ylim=c(0,20),xlab="sum of steps",
                  main="",col=rgb(0,0.4,0,0.35)))
axis(1,at=breaks,labels=ceiling(breaks),las=2,cex.axis=0.75)
mtext("(Before imputation)",side=3)
box()

with(sum.day.imp,hist(sum,breaks=breaks,freq=TRUE,xaxt="n",ylim=c(0,20),xlab="sum of steps",
                  main="",col=rgb(0,0,0.2,0.35)))
axis(1,at=breaks,labels=ceiling(breaks),las=2,cex.axis=0.75)
mtext("(After imputation)",side=3)
box()
title(main="Histogram for the total number of steps taken each day",outer=TRUE)
```

```{r}
name=list(c("mean","median"),c("before imputation","after imputation"))
mean.day.imp=mean(sum.day.imp$sum)
median.day.imp=median(sum.day.imp$sum)
matrix(c(mean.day,median.day,mean.day.imp,median.day.imp),2,2,dimnames=name)
```

**Calculating the average number of steps taken in each interval over weekdays and weekends**
```{r}
data.imp$day=weekdays(data.imp$date,abbreviate=TRUE)

pos=with(data.imp,which(day=="Fri" | day=="Sat" | day=="Sun"))
data.imp$weekfact="weekday"
data.imp$weekfact[pos]="weekend"

sum.week=data.imp %>%
  group_by(weekfact,interval) %>%
  summarise(mean=mean(steps))
```
**Panel plot to compare the avg number of steps taken in each interval over weekday and weekend***
```{r}
library(lattice)
color=c(rgb(0,0.4,0,0.55),rgb(0.4,0,0,0.55))
xyplot(mean~interval | weekfact,data=sum.week,type="l",
       col=color,groups=weekfact,layout=c(1,2),
       xlab="5 minute interval",ylab="mean steps per interval",
       main="Mean steps per interval for weekday and weekend")
```

There is not much significant difference in the walking over weekdays and weekends